---
title: "DKO_ExpressionProfile_010525"
author: "Winnie Lei"
date: "2025-05-01"
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE)
```

## Cholangiocyte expression profile



```{r Packages and Working directory}
setwd(".")
library(NanoStringNCTools)
# library(GeomxTools)
# library(GeoMxWorkflows)
library(knitr)
library(ggplot2)
# library(SpatialDecon)
library(standR)
library(colorspace)
library(ggh4x)
library(tidyverse)
library(ggpubr)
library(rstatix)
library(limma)
library(ExperimentHub)
library(SpatialExperiment)
library(Rtsne)
# library(umap)
# library(SeuratObject)
# library(Seurat)
library(ComplexHeatmap)
# library(nichenetr)
```

## Read in single cell data matrix

Data from https://www.science.org/doi/10.1126/science.aaz6964?url_ver=Z39.88-2003&rfr_id=ori:rid:crossref.org&rfr_dat=cr_pub%20%200pubmed

```{r pressure, echo=FALSE}
Fotis_matrix = Matrix::readMM('../data/Sampaziotis_2021/E-MTAB-8495.aggregated_filtered_normalised_counts.mtx')
library(edgeR)
library(SingleCellExperiment)
library(scran)
sce <- SingleCellExperiment(assays = list(counts = Fotis_matrix))
sce <- computeSumFactors(sce, sizes=seq(20, 80, 5))
sce <- logNormCounts(sce) 
tmm.counts <- assay(sce, "logcounts")
```

```{r Add column and row names}
features <- read.delim("../data/Sampaziotis_2021/E-MTAB-8495.aggregated_filtered_normalised_counts.mtx_rows",
                       header=FALSE)
barcodes <- read.delim("../data/Sampaziotis_2021/E-MTAB-8495.aggregated_filtered_normalised_counts.mtx_cols",
                       header=FALSE)

colnames(tmm.counts) <- barcodes[,1]
rownames(tmm.counts) <- features[,1]
```

```{r Get single cell samples information}
Fotis_sampletable = read.csv("../data/Sampaziotis_2021/E-MTAB-8495.sdrf.txt", sep = "\t")
Fotis_sampletable
```

```{r check samples}
matrix_sample <- stringr::str_extract(barcodes[,1], "[^-]+")
all(matrix_sample %in% Fotis_sampletable$'Comment.BioSD_SAMPLE.')
all(Fotis_sampletable$'Comment.BioSD_SAMPLE.' %in% matrix_sample)
```

## Extract genes of interest:

Signature from email on 2025-03-20:

-General cholangiocyte (CO): KRT-7, KRT-19, GGT, SOX9, HNF1B, CFTR 
-Intrahepatic ducts (IHD): DCDC2, PKHD1, HES4, BICC1, JAG1, SOX4, TACSTD2, HES1 
-Common Bile duct (CBD): TFF3, TFF1, TFF2, AQP5
-Gallbladder (GB): FGF19, SLC15A1, MUC13, SOX17, SLC13A1

```{r create gene list 1}
CO_list_1 = c("KRT7", "KRT19", "GGT1", "SOX9", "HNF1B", "CFTR") # GGT1 instead GGT 
IHD_list_1 = c("DCDC2", "PKHD1", "HES4", "BICC1", "JAG1", "SOX4", "TACSTD2", "HES1") 
CBD_list_1 = c("TFF2", "TFF3", "TFF1","AQP5")
GB_list_1 = c("FGF19", "SLC15A1", "MUC13", "SLC13A1", "SOX17")
```

```{r Convert list to Ensembl}
library("EnsDb.Hsapiens.v75") 
geneIDs2_1 <- ensembldb::select(EnsDb.Hsapiens.v75, 
                              keys= c(CO_list_1,IHD_list_1,CBD_list_1,GB_list_1), 
                              keytype = "SYMBOL", 
                              columns = c("SYMBOL","GENEID"))
geneIDs2_1[geneIDs2_1$SYMBOL=="HNF1B","GENEID"] = "ENSG00000275410"
geneIDs2_1
```

Subset matrix 
```{r Reduce matrix}
Fotis_matrix_reduced_1 = tmm.counts[geneIDs2_1$GENEID[geneIDs2_1$GENEID%in%features[,1]],]
Fotis_matrix_reduced_1 = as.data.frame(as.matrix(Fotis_matrix_reduced_1))
Fotis_matrix_reduced_1
```

Convert geneid to symbol
```{r convert id to symbol}
id2symbol_1 = setNames(as.list(geneIDs2_1$SYMBOL), nm = geneIDs2_1$GENEID)
Fotis_matrix_reduced_1$Symbol = NULL
Fotis_matrix_reduced_1$Symbol = unlist(unname(id2symbol_1[rownames(Fotis_matrix_reduced_1)]))
all(c(CO_list_1,IHD_list_1,CBD_list_1,GB_list_1) %in% Fotis_matrix_reduced_1$Symbol)
```


Remove sparse matrix (more memory)

```{r remove sparse matrix}
rm(Fotis_matrix)
rm(tmm.counts)
gc()
```

## Read DKO data

```{r}
PlateA = readRDS("PlateA_GeoMx.rds")
```

Get relevant dataset
-WT non-humanised: mouse 715TL right
-DKO humanised: mouse 709 NM left and/or 936 NM left; should we stick to one or should we average the expression of all panCK DKO so also including 715 TL left? 

```{r}
PlateA$sampleData$Plate = "PlateA"
all_Plate_Sample = PlateA$sampleData
all_Plate_Sample = cbind(SegmentDisplayName = rownames(all_Plate_Sample), 
                         all_Plate_Sample)
colnames(all_Plate_Sample)[4] = "SlideName"
colnames(all_Plate_Sample)[c(12,13)] = c("ROICoordinateX", "ROICoordinateY")
all_Plate_Sample[all_Plate_Sample$Segment == "panCK","Segment"] = "PanCK"
all_Plate_Sample
```

```{r Get key samples}
Key_Plate_Sample = all_Plate_Sample[(all_Plate_Sample$SlideName == "715 TL right 31")|
                                      (all_Plate_Sample$SlideName == "709 NM left 111")|
                                      (all_Plate_Sample$SlideName == "936 NM left 57"),]
Key_Plate_Sample = Key_Plate_Sample[Key_Plate_Sample$Segment == "PanCK",]
Key_Plate_Sample
```
```{r}
Key_annotation = PlateA$annotationData
Key_annotation = cbind(TargetName = rownames(Key_annotation), 
                             Key_annotation)

Key_Sample_Count = PlateA$data[,rownames(Key_Plate_Sample)]
Key_Sample_Count = cbind(TargetName = rownames(Key_Sample_Count), 
                        as.data.frame(Key_Sample_Count))

key_spe <- standR::readGeoMx(Key_Sample_Count, 
                         Key_Plate_Sample, 
                         featureAnnoFile = Key_annotation,
                         rmNegProbe = FALSE
                         )
key_spe <- geomxNorm(
  key_spe,
  method = "TMM", 
  log = TRUE)

DKO_data = assay(key_spe,"logcounts")
DKO_data = as.data.frame(DKO_data)
# DKO_data = as.data.frame(scale(DKO_data))
```

## Generate heatmap 

```{r Subset DKO data with key genes}
keygenes_1 = c(CO_list_1,IHD_list_1,CBD_list_1,GB_list_1)
keygenes_dko_1 = DKO_data[keygenes_1,]
dim(keygenes_dko_1)
```
```{r Generate heatmap}
keygenes_dko_df_1 = reshape::melt(data.table::data.table(keygenes_dko_1, keep.rownames = TRUE) , id.vars = c("rn"))
colnames(keygenes_dko_df_1)[1] = "gene"
keygenes_dko_df_1 = merge(keygenes_dko_df_1,
                     Key_Plate_Sample[c("Scan Name", "Target", "Sample_ID")],
                     by.x = "variable",
                     by.y = 0,
                     all.x = TRUE)
keygenes_dko_df_1$Sample_ID = sub('.+-(.+)', '\\1', keygenes_dko_df_1$Sample_ID)
keygenes_dko_df_1 = keygenes_dko_df_1[complete.cases(keygenes_dko_df_1), ]
keygenes_dko_df_1$gene = gsub("\\.1","",keygenes_dko_df_1$gene)

# Assigning gene group
keygenes_dko_df_1$Group = ifelse(keygenes_dko_df_1$gene %in% CO_list_1, "General\nCO",
                                         ifelse(keygenes_dko_df_1$gene %in% IHD_list_1, "IHD",
                                                ifelse(keygenes_dko_df_1$gene %in% CBD_list_1, "CBD","GB")))
keygenes_dko_df_1$Group = factor(keygenes_dko_df_1$Group,
                               levels = c("General\nCO", "IHD", "CBD", "GB"))

# Assigning sample group
keygenes_dko_df_1$Target = ifelse(keygenes_dko_df_1$Target == "DKO", "ePCO\nin vivo", "PCO\nin vivo")
keygenes_dko_df_1$Target = factor(keygenes_dko_df_1$Target,
                                levels=c("PCO\nin vivo","ePCO\nin vivo"))

# Changing gene names to match analysis
keygenes_dko_df_1[keygenes_dko_df_1$gene == "GGT1","gene"] <- "GGT"
keygenes_dko_df_1[keygenes_dko_df_1$gene == "CYR61","gene"] <- "CCN1"

```

## Plot general CO only for PCO/ePCO only

```{r Change colours, fig.height = 3, fig.width = 6}
plot_data = keygenes_dko_df_1[!keygenes_dko_df_1$Group == "General\nCO",]
  
plot_data$Target = gsub("ePCO\nin vivo",
                        "atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))", 
                        plot_data$Target)
plot_data$Target = gsub("PCO\nin vivo",
                        "atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))", plot_data$Target)
plot_data$Target = factor(plot_data$Target,
                         levels = c("atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))", 
                                    "atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))"))
  
p <- ggplot(plot_data, aes(Sample_ID, gene, fill = value)) +
         geom_tile(colour = "black")

p + viridis::scale_fill_viridis(direction = 1,
  option = "magma",
  na.value = 'black') +
  facet_grid(Group~Target, scales = "free", space = "free",
             labeller = label_parsed)+ 
  labs(y=NULL, x=NULL, fill="log(TMM)") +
  theme_bw() +
  xlab("") + ylab("") +
  theme(panel.background = element_rect(fill = "black"),
        panel.grid = element_line(color = "black"),
        axis.text.x = element_text(angle = 90, vjust = 0.5, hjust=1),
        strip.background = element_rect(fill="white"),
        strip.text.x = element_text(size = 10,
                                    vjust=-4,
                                    margin = margin(-0.1,0,-0.1,0, "cm")))
```

```{r Store results}
write.csv(plot_data,"../res/SupplFig8I_Data.csv")
```

## Generate dot plot summary

```{r }
DKO_zscore = as.data.frame(scale(DKO_data))
keygenes_zscores = DKO_zscore[keygenes_1,]

keygenes_zscores = reshape::melt(data.table::data.table(keygenes_zscores, keep.rownames = TRUE) , id.vars = c("rn"))
colnames(keygenes_zscores)[1] = "gene"
keygenes_zscores = merge(keygenes_zscores,
                     Key_Plate_Sample[c("Scan Name", "Target", "Sample_ID")],
                     by.x = "variable",
                     by.y = 0,
                     all.x = TRUE)
keygenes_zscores$Sample_ID = sub('.+-(.+)', '\\1', keygenes_zscores$Sample_ID)
keygenes_zscores = keygenes_zscores[complete.cases(keygenes_zscores), ]
keygenes_zscores$gene = gsub("\\.1","",keygenes_zscores$gene)

# Assigning gene group
keygenes_zscores$Group = ifelse(keygenes_zscores$gene %in% CO_list_1, "General\nCO",
                                         ifelse(keygenes_zscores$gene %in% IHD_list_1, "IHD",
                                                ifelse(keygenes_zscores$gene %in% CBD_list_1, "CBD","GB")))
keygenes_zscores$Group = factor(keygenes_zscores$Group,
                               levels = c("General\nCO", "IHD", "CBD", "GB"))

# Assigning sample group
keygenes_zscores$Target = ifelse(keygenes_zscores$Target == "DKO", "ePCO\nin vivo", "PCO\nin vivo")
keygenes_zscores$Target = factor(keygenes_zscores$Target,
                                levels=c("PCO\nin vivo","ePCO\nin vivo"))

# Changing gene names to match analysis
keygenes_zscores[keygenes_zscores$gene == "GGT1","gene"] <- "GGT"
keygenes_zscores[keygenes_zscores$gene == "CYR61","gene"] <- "CCN1"

plot_data = keygenes_zscores[c("Group","Target","value")] %>% 
                          group_by(Group, Target) %>%
                          summarise_all("mean")

plot_data$Target = gsub("ePCO\nin vivo",
                        "atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))", 
                        plot_data$Target)
plot_data$Target = gsub("PCO\nin vivo",
                        "atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))", plot_data$Target)
plot_data$Target = factor(plot_data$Target,
                         levels = c("atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))", 
                                    "atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))"))

ggplot(plot_data,
       aes(x =  Group,
           y = Target,
           color = value,
           size = value)
         ) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "inferno",
                        limits = c(-2,6),
                        name = 'Z-Score') +
  theme(axis.line  = element_blank()) +
  ylab('') +
  xlab('') +
  scale_size(range = c(2,12),
             name = "Z-score",
             limits = c(-2,6))+
  scale_y_discrete(labels = ggplot2:::parse_safe) +
  theme(legend.direction = "vertical", legend.box = "horizontal")
```

```{r Wihtout General CO}
plot_data = plot_data[!plot_data$Group=="General\nCO",]
ggplot(plot_data,
       aes(x =  Group,
           y = Target,
           color = value,
           size = value)
         ) +
  geom_point() +
  theme_bw() +
  scale_color_viridis_c(option = "inferno",
                        limits = c(-2,4),
                        name = 'Z-Score') +
  theme(axis.line  = element_blank()) +
  ylab('') +
  xlab('') +
  scale_size(range = c(4,12),
             breaks = c(-2,0,2,4),
             name = "Z-score",
             limits = c(-2,4))+
  scale_y_discrete(labels = ggplot2:::parse_safe) +
  theme(legend.direction = "vertical", legend.box = "horizontal")
```


## Combine data with single cell

```{r Melt fotis matrix}
Fotis_df_1 = reshape::melt(data.table::data.table(Fotis_matrix_reduced_1, 
                                                keep.rownames = TRUE) , 
                         id.vars = c("rn"))
colnames(Fotis_df_1)[1] = "gene"
Fotis_df_1$variable <- stringr::str_extract(Fotis_df_1$variable, "[^-]+")

Fotis_df_1$value = as.numeric(as.character(Fotis_df_1$value))
Fotis_df_1 = Fotis_df_1[Fotis_df_1$value!=0,]
Fotis_df_1 = Fotis_df_1[complete.cases(Fotis_df_1),]
Fotis_df_1 = as.data.frame(Fotis_df_1 %>%
   group_by(variable, gene) %>% 
   summarise_at(vars("value"), mean))

keygenes_Fotis_df_1 = merge(Fotis_df_1,
                     Fotis_sampletable[c("Comment.BioSD_SAMPLE.", 
                                         "Factor.Value.organism.part.", 
                                         "Factor.Value.growth.condition.")],
                     by.x = "variable",
                     by.y = "Comment.BioSD_SAMPLE.",
                     all.x = TRUE)
keygenes_Fotis_df_1 = unique(keygenes_Fotis_df_1)
keygenes_Fotis_df_1$Gene = unlist(unname(id2symbol_1[keygenes_Fotis_df_1$gene]))

## Add gene group
keygenes_Fotis_df_1$Group = ifelse(keygenes_Fotis_df_1$Gene %in% CO_list_1, "General\nCO",
                                         ifelse(keygenes_Fotis_df_1$Gene %in% IHD_list_1, "IHD",
                                                ifelse(keygenes_Fotis_df_1$Gene %in% CBD_list_1, "CBD","GB")))
keygenes_Fotis_df_1$Group = factor(keygenes_Fotis_df_1$Group,
                                 levels = c("General\nCO",
                                            "IHD",
                                            "CBD",
                                            "GB"))

## Add target group
keygenes_Fotis_df_1$Target = ifelse(keygenes_Fotis_df_1$'Factor.Value.organism.part.' == "common bile duct",
                                  "CBD",
                                  ifelse(keygenes_Fotis_df_1$'Factor.Value.organism.part.' == "gall bladder", 
                                         "GB","IHD"))
keygenes_Fotis_df_1$Target = ifelse(keygenes_Fotis_df_1$'Factor.Value.growth.condition.' == "organoid",
                                  paste0("Org\n",keygenes_Fotis_df_1$Target), paste0("Prim\n",keygenes_Fotis_df_1$Target))
keygenes_Fotis_df_1$Sample_ID = keygenes_Fotis_df_1$variable

# # Add row of FGF19
# fotis_FGF19 = keygenes_Fotis_df[keygenes_Fotis_df$gene == "FGF19",]
# fotis_FGF19$Group = "GB"

# Changing gene names to match analysis
keygenes_Fotis_df_1[keygenes_Fotis_df_1$Gene == "GGT1","Gene"] <- "GGT"
keygenes_Fotis_df_1[keygenes_Fotis_df_1$Gene == "CYR61","Gene"] <- "CCN1"

# keygenes_Fotis_df = rbind(keygenes_Fotis_df,fotis_FGF19)

```

WT, DKO, Primary IHD, Primary CBD, Primary GB, Organoid IHD, Organoid CBD, Organoid GB. 

```{r Combine data}
keygenes_Fotis_df_1$gene = keygenes_Fotis_df_1$Gene
all_data_1 = rbind(keygenes_Fotis_df_1[c("Sample_ID","gene","value","Target","Group")],
                 keygenes_dko_df_1[c("Sample_ID","gene","value","Target","Group")]
)

## make italic
# all_data_1$Target = gsub("in vivo", expression(italic("in vivo")), all_data_1$Target)
  
all_data_1$Target = factor(all_data_1$Target,
                         levels = c("ePCO\nin vivo",
                                    "PCO\nin vivo",
                                    "Org\nIHD",
                                    "Org\nCBD",
                                    "Org\nGB",
                                    "Prim\nIHD",
                                    "Prim\nCBD",
                                    "Prim\nGB"))

```



## Grouped barplot with signature 1

```{r Grouped bar plot, fig.width = 10, fig.height = 4}
bar_all_data_1 = all_data_1
bar_all_data_1 = bar_all_data_1[!grepl("Org",bar_all_data_1$Target),]

bar_all_data_1$Target = gsub("ePCO\nin vivo",
                        "atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))", 
                        bar_all_data_1$Target)
bar_all_data_1$Target = gsub("PCO\nin vivo",
                        "atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))", bar_all_data_1$Target)
bar_all_data_1$Target = gsub("Prim\nIHD",
                        "atop(atop(textstyle('Prim'),textstyle('IHD')))", bar_all_data_1$Target)
bar_all_data_1$Target = gsub("Prim\nCBD",
                        "atop(atop(textstyle('Prim'),textstyle('CBD')))", bar_all_data_1$Target)
bar_all_data_1$Target = gsub("Prim\nGB",
                        "atop(atop(textstyle('PCO'),textstyle('GB')))", bar_all_data_1$Target)
bar_all_data_1$Target = factor(bar_all_data_1$Target,
                         levels = c("atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))", 
                                    "atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))",
                                    "atop(atop(textstyle('Prim'),textstyle('IHD')))",
                                    "atop(atop(textstyle('Prim'),textstyle('CBD')))",
                                    "atop(atop(textstyle('PCO'),textstyle('GB')))"))

```


```{r plot General CO signature only}
bar_all_data_wo_CO = bar_all_data_1[bar_all_data_1$Group == "General\nCO",]

ggplot(bar_all_data_wo_CO, aes( y = value, x = Target), fill = "#7D7D7D", color = "#7D7D7D") +
  geom_bar(
    position = "dodge", width = 0.8, stat = "summary", fun = "mean",
    color = "#7D7D7D", fill = "#7D7D7D", linewidth = .8
  ) +
  geom_point(colour = "black", fill = "white",
    alpha = 0.3
  ) + 
  # stat_summary(
  #   fun.data = mean_se, geom = "errorbar", color = "black",
  #   position = position_dodge(0.8), width = 0.2, linewidth = 0.8,  na.rm = TRUE
  # )  +
  stat_summary(fun.ymin = function(x) mean(x) - sd(x), 
           fun.ymax = function(x) mean(x) + sd(x), position ='dodge', 
           geom = 'errorbar', color = "black",
            width = 0.2, linewidth = 0.8,  na.rm = TRUE) + 
  theme_classic() +
  theme(legend.position = "None") +
  labs(y="Mean Expresion (log(TMM))", x=NULL, fill="#7D7D7D") +
  theme(aspect.ratio = 5/4) +
      # stat_summary(aes(label=round(..y..,2)), fun.y=mean, geom="text", size=4,
      #        vjust = -6,  na.rm = TRUE) +
  scale_y_continuous(expand =  expansion(mult = c(0, 0), 
                                         add = c(0, 1)),
                     limits = c(0, 10)) +
  scale_x_discrete(labels = parse(text = levels(bar_all_data_wo_CO$Target))) 
```


```{r Save barplot data}
write.csv(bar_all_data_wo_CO,"../res/BarPlot_Data.csv")
```

## Principal component analysis

```{r Perform PCA for signature data}
library(factoextra)

signature_df = reshape(all_data_1[c("Sample_ID","gene","value")], idvar = "Sample_ID", timevar = "gene", direction = "wide")
rownames(signature_df) = signature_df$Sample_ID
signature_df$Sample_ID = NULL
signature_df[is.na(signature_df)] <- 0
res.pca <- prcomp(signature_df, scale = TRUE)
fviz_eig(res.pca)

groups <- as.factor(unique(all_data_1[c("Sample_ID","Target")])$Target)
fviz_pca_ind(res.pca,
             col.ind = groups, # color by groups
             palette = RColorBrewer::brewer.pal(n = 8,"Dark2"),
             addEllipses = TRUE, # Concentration ellipses
             ellipse.type = "confidence",
             legend.title = "Groups",
             repel = TRUE
             )
```

```{r Repeat without Organoid data}
all_data_1_woOrg = all_data_1[!grepl("Org",all_data_1$Target),]
all_data_1_woOrg$Target = factor(all_data_1_woOrg$Target,
                                 levels = c("PCO\nin vivo",
                                            "ePCO\nin vivo",
                                            "Prim\nIHD",
                                            "Prim\nCBD", 
                                            "Prim\nGB"))
signature_df = reshape(all_data_1_woOrg[c("Sample_ID","gene","value")], idvar = "Sample_ID", timevar = "gene", direction = "wide")
rownames(signature_df) = signature_df$Sample_ID
signature_df$Sample_ID = NULL
signature_df[is.na(signature_df)] <- 0
```

```{r Repeat without Organoid data or general CO}
all_data_1_woOrg$Target = factor(all_data_1_woOrg$Target,
                                 levels = c("PCO\nin vivo",
                                            "ePCO\nin vivo",
                                            "Prim\nIHD",
                                            "Prim\nCBD", 
                                            "Prim\nGB"))
plot_data = all_data_1_woOrg[!all_data_1_woOrg$Group=="General\nCO",]

signature_df = reshape(plot_data[c("Sample_ID","gene","value")], idvar = "Sample_ID", timevar = "gene", direction = "wide")
rownames(signature_df) = signature_df$Sample_ID
signature_df$Sample_ID = NULL
signature_df[is.na(signature_df)] <- 0
res.pca <- prcomp(signature_df, scale = TRUE)
fviz_eig(res.pca)

groups <- as.factor(unique(plot_data[c("Sample_ID","Target")])$Target)
fviz(res.pca, 
     element="ind", 
     geom="point",  
     pointsize = 3,
     habillage=groups,
     addEllipses=TRUE, 
     ellipse.type = "confidence",
     palette = RColorBrewer::brewer.pal(n = 8,"Dark2"), 
     invisible="quali")  + 
  labs(title = "", x = "PC1", y = "PC2")
```
```{r Prettify PCA plots}
pca_res = res.pca$x

group_dict = list("PCO\nin vivo"="atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))",
                  "ePCO\nin vivo"="atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))",
                  "Prim\nIHD"="atop(atop(textstyle('Prim'),textstyle('IHD')))",
                  "Prim\nCBD"= "atop(atop(textstyle('Prim'),textstyle('CBD')))",
                  "Prim\nGB"="atop(atop(textstyle('Prim'),textstyle('GB')))")

groups <- factor(unique(plot_data[c("Sample_ID","Target")])$Target)

groups = factor(unlist(unname(group_dict[groups])),
                levels = c("atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))",
                 "atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))",
                 "atop(atop(textstyle('Prim'),textstyle('IHD')))",
                 "atop(atop(textstyle('Prim'),textstyle('CBD')))",
                 "atop(atop(textstyle('Prim'),textstyle('GB')))"))
legend_group = c("atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))",
                 "atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))",
                 "atop(atop(textstyle('Prim'),textstyle('IHD')))",
                 "atop(atop(textstyle('Prim'),textstyle('CBD')))",
                 "atop(atop(textstyle('Prim'),textstyle('GB')))")

batch = ifelse(grepl("PCO",groups),"ePCOs","Primary")

batch_group = c("atop(atop(textstyle('e/PCOs'),textstyle(italic(''))))",
               "atop(atop(textstyle('Sampaziotis'),textstyle(italic('et al'))))")

pca_res %>% 
  ggplot(aes(x=PC1, y=PC2, fill = groups, shape = batch))+
  geom_point(aes(fill=groups,
                 shape=factor(batch)), 
             size=4,
             stroke =1)+
  ggforce::geom_mark_ellipse(aes(fill = groups))+
  scale_fill_brewer(palette="Dark2",
                    labels = parse(text = legend_group),
                    name="Sample\nGroup" )+ 
  scale_shape_manual(values = c("ePCOs" = 21,
                                "Primary" = 22),
                    labels = parse(text = batch_group),
                    name="Batch") +
  theme_bw() + 
  theme(legend.text = element_text(vjust =-4),
        legend.text.align = 0.5,
        legend.title.align = 0.5)
```
```{r Save PCA results}
write.csv(pca_res,"../res/PCA_results.csv")
```

```{r Correlations between signature data without General CO}
library(corrplot)
plot_data = all_data_1_woOrg[!all_data_1_woOrg$Group=="General\nCO",]
signature_df = reshape(plot_data[c("Sample_ID","gene","value")], idvar = "Sample_ID", timevar = "gene", direction = "wide")
rownames(signature_df) = signature_df$Sample_ID
signature_df$Sample_ID = NULL
signature_df[is.na(signature_df)] <- 0
signature_df = t(signature_df)

res = cor(signature_df)
corrplot(res, order = "hclust", 
         addrect = 2,
         method = "color",
         tl.col = "black", tl.srt = 90,tl.cex = 0.5,
         col = rev(COL2('RdBu', 10)))
```

```{r Generate average }
library(data.table)
correlation_tbl= reshape2::melt(res)

sample_dict = unique(all_data_1_woOrg[c("Sample_ID","Target")])
sample_dict = setNames(sample_dict$Target,sample_dict$Sample_ID)

correlation_tbl$Var1_group = sample_dict[correlation_tbl$Var1]
correlation_tbl$Var2_group = sample_dict[correlation_tbl$Var2]

correlation_tbl_1 = correlation_tbl %>%
     group_by(Var1_group, Var2_group) %>%
     summarise(across(value, mean))

correlation_tbl_1 = as.data.frame(correlation_tbl_1)
correlation_tbl_1[correlation_tbl_1$Var1_group == correlation_tbl_1$Var2_group,"value"] = 1

correlation_tbl_1 = reshape(correlation_tbl_1, 
                            idvar="Var1_group", 
                            timevar="Var2_group", 
                            direction="wide")

n_labels = c("atop(atop(textstyle('ePCO'),textstyle(italic('in vivo'))))",
                 "atop(atop(textstyle('PCO'),textstyle(italic('in vivo'))))",
                 "atop(atop(textstyle('Prim'),textstyle('IHD')))",
                 "atop(atop(textstyle('Prim'),textstyle('CBD')))",
                 "atop(atop(textstyle('Prim'),textstyle('GB')))")
rownames(correlation_tbl_1) = correlation_tbl_1$Var1_group
correlation_tbl_1$Var1_group = NULL
colnames(correlation_tbl_1) = gsub("value.", "",colnames(correlation_tbl_1))
ggcorrplot::ggcorrplot(correlation_tbl_1, hc.order = FALSE,
   outline.color = "black",
   ggtheme = ggplot2::theme_bw,
   colors = RColorBrewer::brewer.pal(n = 9, name = "RdBu")[c(9,5,1)],
   lab = FALSE,
   legend.title = "Expression\nCorrelations") + 
  theme(axis.text.x = element_text(angle = 0, vjust = 0.75, hjust = 0.5),
        axis.text.y = element_text(vjust = 0.75)) + 
  scale_x_discrete(labels= parse(text = (n_labels))) +
  scale_y_discrete(labels= parse(text = (n_labels)))

```

```{r Save data}
write.csv(correlation_tbl_1,"../res/CorrelationMatrix.csv")
```


Session information
```{r}
sessionInfo()
```



